The formal verification of the ctm approach to forcing \thanks{Supported by Secyt-UNC project 33620180100465CB and Conicet.}

### 2022-09-18 Independence_CH/sample_proof.tex
More details about Naproche

Undoubtedly, even for this cherry-picked example, the formalization
looks a bit codish. It is therefore inevitable to compare this to the
magnificent results obtained by P.~Koepke and his team by using
Isabelle/Naproche \cite{10.1007/978-3-030-81097-9_2} (particularly,
the proof of König's Theorem). The trick there consists in presenting the result
being formalized as a restricted first-order problem, and then each
proof step can be handled by an automatic theorem prover.

>>>>

This cherry-picked example shows that the formalization can be close
to the mathematical exposition and might be useful to reconstruct the
proof from the book; nonetheless, it also has significantly more
details than the mathematical prose, even with some indications to
direct the automatic tools.

There has been some progress on assistants where one writes statements
and proofs in natural language; recently P.~Koepke and his team
achieved magnificient results by using Isabelle/Naproche
\cite{10.1007/978-3-030-81097-9_2} to formalize proofs of several
results (particularly, the proof of König's Theorem). The input
language of Isabelle/Naproche is a \emph{controlled} natural language
that presents the result being formalized as a deduction in
first-order logic, where every assumption and the ``whole logical
scenario'' are explicitly given. From the input language,
Isabelle/Naproche builds ``proof tasks'' that are handled to automatic
theorem provers. As far as we can tell, Isabelle/Naproche is
promising but still unsuitable for a project of the magnitude of ours.

% Undoubtedly, even for this cherry-picked example, the formalization
% looks a bit codish. It is therefore inevitable to compare this to the
% magnificent results obtained by P.~Koepke and his team by using
% Isabelle/Naproche \cite{10.1007/978-3-030-81097-9_2} (particularly,
% the proof of König's Theorem). The trick there consists in presenting the result
% being formalized as a restricted first-order problem, and then each
% proof step can be handled by an automatic theorem prover.
